---
title : "Time Series part 2"
output: github_document
---
---
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
data(AirPassengers)
AirPassengers
```
```{r}
plot(AirPassengers)
```
> It shows trends and seasonality

Auto-corellation plot

```{r}
tmp=acf(AirPassengers,type="cor",plot = FALSE)
tmp$acf[1:3,1,1]
plot(tmp)
```
Differencing

```{r}
log_AirPassengers = log(AirPassengers)
diff_serie = diff(log_AirPassengers,lag=12,difference=1)
diff_serie2 = diff(log_AirPassengers,lag=12,difference=2)

par(mfrow=c(3,1))
plot(AirPassengers)
plot(diff_serie)
plot(diff_serie2)
```

> The differenciated series does not seem stationary

Is it a white noise?

```{r}
Box.test(diff_serie,lag=10,type="Ljung-Box")
```

> the diff-serie can be assimilated to a white noise.

ACF on the residual serie

```{r}
tmp=acf(diff_serie,type="cor",plot = FALSE)
tmp2=acf(diff_serie2,type="cor",plot = FALSE)
par(mfrow=c(2,1))
plot(tmp)
plot(tmp2)
```

> Almost the same plot where difference = 1 or 2.




#### Example with Google stock price ####

```{r}
library(ggplot2)
library(fpp2)
ggtsdisplay(goog200)
```
> We see only a trend. No seosonal patern


Differencing

```{r}
diff_serie = diff(goog)
ggtsdisplay(diff_serie)
```

> Auto corellation: 95% of the time the corelation should be inside the band. It is ok to have only a peak (= the 5%)

White Noise

```{r}
Box.test(diff_serie,lag=25,type="Ljung-Box")
```

> Large p-value => We have a white noise => Nothing to forecast



##### ARMA #####

Function arima.sim allows to simulate an ARp.

```{r}
modele<-list(ar=c(0.8))
ar1<-arima.sim(modele,1000)

par(mfrow=c(2,2))
plot.ts(ar1)
acf(ar1)
pacf(ar1)
```

Function arima.sim allows to simulate an MAq

```{r}
modele<-list(ma=c(0.8))
ma1<-arima.sim(modele,1000)

par(mfrow=c(2,2))
plot.ts(ma1)
acf(ma1)
pacf(ma1)
```







Exercices: Income, Unemployment

1- Plot the data. Identify unusual informations. Understand patterns
2- If necessary, use a Box-Cox transformation (logarithm or power for example)
3- If necessary, difference the data until it appears stationary
4- Plot the ACF/PACF of the differenced data and try to determine possible candidate models
5- Try your chosen models(s) and use the AICc to serach for a better model
6- Check the residuals from your chosen model by plotting the ACF of the residuals, and doing a Box test of the residuals
7- If residuals look like white noise calculate forecast. If not, go back to 4-

```{r}
library(fpp2)
autoplot(uschange[,c("Income","Unemployment")])
```
> No need to transform the data (since there is no evidence of changing variance).
> No need to differentiate the series because there already appears to be stationary.

Box test to check the check if the series are just white noise or not
```{r}
Box.test(uschange[,c("Income")],lag=10,type="Ljung-Box")
Box.test(uschange[,c("Unemployment")],lag=10,type="Ljung-Box")
```

> For the first dataset p-value is high => white noise => Nothing to forecast, the best forecast is SES = "mean" in this case (no trend, no season)
For the second dataset p-value is low => no white noise => we can go further.

ACF and PACF plot to identify possible model ARp or MAq 

```{r}
par(mfrow=c(2,1))
ggtsdisplay(uschange[,"Income"])
ggtsdisplay(uschange[,"Unemployment"])
```

> The PACF/ACF graphs are suggesting to try first a AR8 model.

ARIMA(8,0,0)

```{r}
AR8 = Arima(uschange[,"Unemployment"],order=c(8,0,0))
AR8
```

Plot of the residuals for the ARIMA(8,0,0) model

```{r}
checkresiduals(AR8)
```

> The ACF plot of the residuals from the ARIMA(8,0,0) model shows that all autocorellations are within the threshold limits indicating that the residuals are behaving like white noise. The large p-value reported by the Ljung-Box test also suggests the same.


Forecast from the chosen model

```{r}
autoplot(forecast(AR8))
```

#### Exercice: Production #### 

```{r}
library(fpp2)
Production = uschange[,c("Production")]
autoplot(Production)
```
> No need to transform the data (since there is no evidence of changing variance).

```{r}
ggtsdisplay(Production)
```
> The PACF/ACF graphs are suggesting to try first a AR4, AR12 or MA8 model.

```{r}
Box.test(Production,lag=10,type="Ljung-Box")
```

> p-value is low => no white noise => we can go further


```{r}
AR4 = Arima(Production,order=c(4,0,0))
AR12 = Arima(Production,order=c(12,0,0))
AR4
AR12
```

```{r}
checkresiduals(AR4)
checkresiduals(AR12)
```

> The ACF plot of the residuals from the ARIMA(12,0,0) model shows that all autocorellations are within the threshold limits indicating that the residuals are behaving like white noise. The large p-value reported by the Ljung-Box test also suggests the same.
For the ARIMA(4,0,0) there is a corellation with Lag 8 and the p-value is rather low.
=> Best model for forecast = ARIMA(12,0,0)

```{r}
autoplot(forecast(AR12))
```


#### Exercice: Google #### 

```{r}
library(fpp2)
ggtsdisplay(goog)
```
> No need to transform the data (since there is no evidence of changing variance).
> Since there is a trend in the data (confirmed by ACF plot) let's take a first difference of the data.)

```{r}
diff2_goog = diff(goog, difference=2)
ggtsdisplay(diff2_goog)
```
> The differenced data look stationary and so we will not consider further difference.
> The PACF/ACF graphs are suggesting to try ARIMA(0,2,1) or ARIMA(0,2,28) model.

```{r}
Box.test(goog,lag=10,type="Ljung-Box")
```

> p-value is low => no white noise => we can go further


```{r}
Arima_fit = Arima(goog,order=c(0,2,28))
Arima_fit2 = Arima(goog,order=c(2,2,28))
Arima_fit
Arima_fit2
```

```{r}
checkresiduals(Arima_fit)
checkresiduals(Arima_fit2)
```

> The ACF plot of the residuals from the ARIMA(12,0,0) model shows that all autocorellations are within the threshold limits indicating that the residuals are behaving like white noise. The large p-value reported by the Ljung-Box test also suggests the same.
For the ARIMA(4,0,0) there is a corellation with Lag 8 and the p-value is rather low.
=> Best model for forecast = ARIMA(12,0,0)

```{r}
autoplot(forecast(Arima_fit))
```

Split between train and test sets (last 5 days)

```{r}
goog_train<-head(goog, length(goog)-6)
goog_test<-tail(goog, 5)

plot(goog_train)
lines(goog_test,col=2)
```

```{r}
Arima_fit_train = Arima(goog_train,order=c(0,2,28))
autoplot(goog_train) + 
autolayer(goog_test) +
autolayer(forecast(Arima_fit_train), series='Arima_fit', PI=FALSE)
```






#### Exercice: Covid #### 

```{r}
covid19 = read.csv('C:/Users/jakam/Documents/R/datasets/WHO-COVID-19-global-data.csv')
covid19_F=covid19[covid19$Country=="France",]
covid19_F_nc=ts(covid19_F$New_cases)
library(forecast)
library(ggplot2)
autoplot(covid19_F_nc) +
ggtitle('Number of new cases of Covid19 in France')+ xlab('days')+
ylab('Number of new cases of Covid19')

```

> No need to transform the data (since there is no evidence of changing variance).
> Since there is a trend in the data (confirmed by ACF plot) let's take a first difference of the data.)

```{r}
diff_covid19_F_nc = diff(covid19_F_nc, difference=3)
ggtsdisplay(diff_covid19_F_nc)
```
> The PACF/ACF graphs are suggesting to try first a ARIMA(0,2,14) model.

```{r}
Box.test(covid19_F_nc,lag=10,type="Ljung-Box")
```

> p-value is low => no white noise => we can go further


```{r}
Arima_fit = Arima(covid19_F_nc,order=c(1,2,19))
Arima_fit
```

```{r}
checkresiduals(Arima_fit)
```

> The ACF plot of the residuals from the ARIMA(12,0,0) model shows that all autocorellations are within the threshold limits indicating that the residuals are behaving like white noise. The large p-value reported by the Ljung-Box test also suggests the same.
For the ARIMA(4,0,0) there is a corellation with Lag 8 and the p-value is rather low.
=> Best model for forecast = ARIMA(12,0,0)

```{r}
autoplot(forecast(Arima_fit))
```
